## Reproducible Research
First create a directory to store the dataset
```{r createdirectory, echo=TRUE}
if(!file.exists("data")) {
        dir.create("data")
}
```
Download the dataset
```{r downloadfile, echo = TRUE}
downloadUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(downloadUrl,"./data/projectDataset.zip")
```

Extract the dataset
```{r extract, echo=TRUE}
unzip("./data/projectDataset.zip")
```

Install lsit of packages iof not installed
```{r installpackages, echo=TRUE}
list.of.packages <- c("knitr","ggplot2","timeDate")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if (length(new.packages))
        install.packages(new.packages)
```
Read the file
```{r readfile, echo=TRUE}
activity <- read.csv("activity.csv") ;str(activity)
```
Format date and time columns
```{r formatdatetime, echo=TRUE}
activity$date <- as.Date(activity$date)
starttime <- as.POSIXct("2012-10-01 00:00:00")
interval_seq <- seq(starttime, length = nrow(activity), by = 300) #5 min interval or 300 secs
activity <- cbind(activity,interval_seq)
```

Get number of steps taken each day

```{r activitysum, echo=TRUE }
activity_sum <- aggregate(activity$steps,by = list(activity$date), FUN = "sum")
allNAs <- complete.cases(activity)
allNAs <- activity[allNAs,]
```
Plot the result
```{r plotsum, fig.height=6}
library(ggplot2)
str(activity_sum)
qplot(x, data = activity_sum, geom = "histogram",xlab = "steps", main = "Histogram of Total Number of Steps"
)
```
  
Get the mean and median of total steps

```{r getmeanmedian, echo=TRUE}
 mean(activity_sum$x, na.rm = TRUE)
 median(activity_sum$x, na.rm = TRUE)
```
Get the activity mean
```{r getmean, echo=TRUE}
activity_mean <- aggregate(
        activity$steps,by = list(activity$date), FUN = "mean", simplify  = TRUE
)
```
Get activity sequence
```{r getsequence, echo=TRUE}
activity_seq <- activity$interval_seq[1:nrow(activity_mean)]
plotactivity <- cbind(activity_mean,activity_seq)
good <- complete.cases(plotactivity)
plotactivity_good <- plotactivity[good,]
```
Plot the result
```{r plot2, fig.height=6}
qplot(
        activity_seq,x, data = plotactivity_good, geom = "line",colour = x, ylab = "Avg steps across all days"
        , xlab = "5-Minute Interval", main = "Average Daily Activity Pattern"
)
```
  
Which 5-minute interval, on average across all the days in 
the dataset, contains the maximum number of steps
```{r maxsteps, echo=TRUE}
plotactivity$Group.1[which.max(plotactivity$x)]
```
Number of NAs
```{r numberofNAs, echo=TRUE}
nrow(activity[!complete.cases(activity),])
```
I will replace with the mean of the 5minute interval
```{r meaninterval,echo=TRUE}
library(dplyr)
mean_by_interval <- activity %>%
        group_by(date) %>%
        summarise_each(funs(mean))
#the mean of the 5-minute interval by dates are all the same at 1177.5
activity$steps[which(is.na(activity$steps))] <- 1177.5
```
Create a new dataset from here with the NAs filled in
```{r createataset,echo=TRUE}
write.csv(activity,file = "activity_with_NAs_filled.csv")
```
total number of steps taken each day with NA filled with plot

```{r totalsteps, echo=TRUE}
activity_sum_noNA <- aggregate(activity$steps,by = list(activity$date), FUN = "sum")

qplot(
        log(x), data = activity_sum_noNA, geom = "histogram",xlab = "steps", main =
                "Histogram of log of Total Number of Steps NAs filled",binwidth = 0.1
)
```
  
mean and median with NAs filled
```{r meanmedian1, echo=TRUE}
 mean(activity_sum_noNA$x, na.rm = TRUE)
 median(activity_sum_noNA$x, na.rm = TRUE)
```
Make a factor of weekdays and weekends
```{r weeksfactor, echo=TRUE}
library(timeDate)
new_activity <- activity
new_activity$week_days <-  ifelse(isWeekday(activity$date, wday = 1:5), "Weekday","Weekend")
new_activity$week_days <- factor(new_activity$week_days)
str(new_activity)
week_days <- new_activity[new_activity$week_days == "Weekday",]
week_ends <- new_activity[new_activity$week_days == "Weekend",]

wkdayaveragesteps <- aggregate(week_days$steps,by = list(week_days$date), FUN = "mean")

starttime <- as.POSIXct("2012-10-01 00:00:00")
interval_seq <- seq(starttime, length = nrow(wkdayaveragesteps), by = 300) #5 min interval or 300 secs
wkdayavg<- cbind(wkdayaveragesteps,interval_seq)

wkendaveragesteps <- aggregate(week_ends$steps,by = list(week_ends$date), FUN = "mean")

starttime <- as.POSIXct("2012-10-01 00:00:00")
interval_seq <- seq(starttime, length = nrow(wkendaveragesteps), by = 300) #5 min interval or 300 secs
wkendavg<- cbind(wkendaveragesteps,interval_seq)
```

Plot the weekdays and weekends
```{r plotweeks,fig.height=6}
qplot(
        interval_seq,log(x), data = wkdayavg, geom = "line", ylab = "Avg steps across all days"
        , xlab = "5-Minute Interval", main = "Weekday Average Daily Activity Pattern"
)

qplot(
        interval_seq,log(x), data = wkendavg, geom = "line", ylab = "Avg steps across all days"
        , xlab = "5-Minute Interval", main = "Weekend Average Daily Activity Pattern"
)
```